---
title: "R Notebook"
output: html_notebook
---


```{r}
# BIKE SHARE DEMAND
bike <- read.csv('C:/Users/Jai Katariya/Desktop/Jai Katariya/Self Learning/R/Course 2-R-Course-HTML-Notes/R-Course-HTML-Notes/R-for-Data-Science-and-Machine-Learning/Training Exercises/Machine Learning Projects/CSV files for ML Projects/bikeshare.csv')
bike <- as.data.frame(bike)
head(bike)
```
```{r}
str(bike)
```

```{r}
summary(bike)
```

```{r}
tail(bike)
```

Exploratory Data Analysis¶
```{r}

library(Amelia)
missmap(bike, main = 'missing map', col = c('yellow', 'black'), legend =F)

```

 Create a scatter plot of count vs temp.
 
```{r}

library(ggplot2)

temp_scatterplot <- ggplot(bike, aes(temp, count)) +
  
                    geom_point(aes(color = temp),alpha = 0.2)+ 
  
                    xlab('Temperature') + ylab("Count of Rentals")

print(temp_scatterplot)

```
Plot count versus datetime as a scatterplot with a color gradient based on temperature. 
Convert the datetime column into POSIXct before plotting.

```{r}

bike$datetime <- as.POSIXct(bike$datetime)

datetime_scatterplot <- ggplot(bike, aes(datetime, count)) + 
  
                        geom_point(alpha = 0.3, aes(color = temp))+
  
                        scale_color_gradient(low = 'green', high = 'red') +
  
                        xlab('Time') + ylab("Count of Rentals")

print(datetime_scatterplot)

```
```{r}

cor_tempVscount <- cor(bike[, c('temp', 'count')])
print(cor_tempVscount)

```

Created a boxplot to explore the season data with the y axis indicating count and the x axis begin a box for each season.

```{r}

seasons <- ggplot(bike, aes(factor(season), count)) + geom_boxplot(aes(color = factor(season))) + xlab('Season') + ylab("Count of Rentals")

print(seasons)

```

Feature Engineering

Before dealing with date time column, we need to feature it.

Created an "hour" column that takes the hour from the datetime column.

```{r}

bike$hour <- sapply(bike$datetime, function(x){format(x, '%H')})
head(bike)
```
Now create a scatterplot of count versus hour, with color scale based on temp. Only use bike data where workingday==1.
Additions:Used the additional layer: scale_color_gradientn(colors=c('color1',color2,etc..)) where the colors argument is a vector gradient of colors you choose, not just high and low. 
Used position=position_jitter(w=1, h=0) inside of geom_point() and check out what it does.

```{r}

bike_data <- subset(bike, bike$workingday == 1)

hour_scatterplot <- ggplot(bike_data,aes(hour, count)) + geom_point(aes(color = temp), 
                    
                    position =  position_jitter(w = 1, h=0), alpha = 0.5)+ scale_color_gradientn(colors= c('dark blue', 'blue', 'light green', 'yellow', 'orange', 'red')) + xlab('Hour') + ylab("Count of Rentals")

print(hour_scatterplot)

```

```{r}

library(dplyr)
hour2_scatterplot <- ggplot(filter(bike, workingday ==0), aes(hour, count)) + 
  
                    geom_point(aes(color = temp), position = position_jitter(w =1, h = 0), 
                               
                    alpha = 0.5) + scale_color_gradientn(colors= c('dark blue', 'blue', 
                                                                   
                    'light green', 'yellow', 'orange', 'red')) + xlab('Hour') + ylab("Count of Rentals")

print(hour2_scatterplot)

```

Compare Count of Rentals with Casual and Registered.

```{r}

Ca_scatterplot <- ggplot(bike,aes(hour, casual)) + geom_point(colour ='Red') + xlab('Hour') + ylab("Count of Casual Rentals")
print(Ca_scatterplot)

```
```{r}

Re_scatterplot <- ggplot(bike,aes(hour, registered)) + geom_point(colour ='blue') + xlab('Hour') + ylab("Count of Registered Rentals")
print(Re_scatterplot)

```

For an interactive plot, I installed plotly library. 

```{r}
#install.packages('plotly')
library(ggplot2)
library(plotly)

CountVsTemp <- ggplot(bike, aes(temp, count)) + geom_point(aes(color = temp)) + xlab("Temperature") + ylab("Count of Rentals")

gpl <- ggplotly(CountVsTemp)

print(gpl)
```

